---
title: "Calculate biodiversity indicator uncertainty"
subtitle: "Bootstrapping"
author: "Ward Langeraert"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load packages
library(b3gbi)     # Biodiversity indicators for data cubes
library(tidyverse) # Data wrangling and visualisation
library(here)      # Relative paths

# Source functions
source(here("source", "R", "evenness_formula.R"))
source(here("source", "R", "boot_statistic.R"))
source(here("source", "R", "perform_bootstrap.R"))
source(here("source", "R", "bootstrap_list_to_df.R"))
source(here("source", "R", "get_bootstrap_ci.R"))
```

# Exploration of the b3gbi package

We install b3gbi package version 0.2.1.
We load a dataset from the package.
How does the cube look like?

```{r}
# Load GBIF data cube
cube_name <- system.file("extdata", "europe_insect_cube.csv", package = "b3gbi")

# Load taxonomic info for cube
tax_info <- system.file("extdata", "europe_insect_info.csv", package = "b3gbi")

insect_data_df <- read.csv(cube_name)
head(insect_data_df)
```

We process the cube.

```{r}
# Prepare cube
insect_data <- process_cube_old(cube_name, tax_info)

insect_data
```

> Where did the info on min. coordinate uncertainty in meters go?

```{r}
# Calculate diversity metric
map_obs_rich_insects <- obs_richness_map(insect_data)

print(map_obs_rich_insects)
```

```{r}
# Plot diversity metric
plot(map_obs_rich_insects,
     title = "Observed Species Richness: Insects in Europe")
```

## Calculate an index over time

Let's calculate Pietlou's eveness over time using b3gbi.

```{r}
eveness_insect_data <- pielou_evenness_ts(insect_data)
head(eveness_insect_data)
```

```{r}
plot(eveness_insect_data)
```

> Visualisation of trend not correct.

Let's calculate this metric from scratch.
We select the data after 2010 to ensure calculations with enough data.

```{r}
insect_data_new <- insect_data$data %>%
  group_by(year) %>%
  mutate(n = sum(obs)) %>%
  ungroup() %>%
  filter(year > 2010) %>%
  select(-n)

insect_data_new %>%
  group_by(year) %>%
  summarise(n_obs = sum(obs)) %>%
  ggplot() +
    geom_bar(aes(x = as.factor(year), y = n_obs), stat = "identity") +
    labs(x = "year", y = "number of observations")
```

## Bootstrapping

We perform bootstrapping for evenness (`boot::boot()`).
Bootstrapping is a statistical technique used to estimate the distribution of a statistic by resampling with replacement from the original data.
Below are the common notations used in bootstrapping:

1. **Original Sample Data**: $\mathbf{X} = \{X_1, X_2, \ldots, X_n\}$
    - The initial set of observed data points. Here, $n$ is the sample size.

2. **Bootstrap Sample**: $\mathbf{X}^* = \{X_1^*, X_2^*, \ldots, X_n^*\}$
    - A sample of size $n$ drawn with replacement from the original sample $\mathbf{X}$. Each $X_i^*$ is drawn independently from $\mathbf{X}$.

3. **Statistic of Interest**: $\theta$
    - The parameter or statistic being estimated, such as the mean $\bar{X}$, variance $\sigma^2$, or median.

4. **Bootstrap Replication**: $\hat{\theta}^*_b$
    - The value of the statistic of interest calculated from the $b$-th bootstrap sample $\mathbf{X}^*_b$. For example, if $\theta$ is the sample mean, $\hat{\theta}^*_b = \bar{X}^*_b$.

5. **Number of Bootstrap Samples**: $B$
    - The number of bootstrap samples drawn from the original data. Common choices for $B$ are 1000 or 10,000 to ensure a good approximation of the distribution.

6. **Bootstrap Estimate of the Statistic**: $\hat{\theta}_{\text{boot}}$
    - The average of the bootstrap replications:

$$
\hat{\theta}_{\text{boot}} = \frac{1}{B} \sum_{b=1}^B \hat{\theta}^*_b
$$

7. **Bootstrap Standard Error**: $\text{SE}_{\text{boot}}$
    - The standard deviation of the bootstrap replications, which estimates the variability of the statistic.

8. **Bootstrap Bias**: $\text{Bias}_{\text{boot}}$
    - This bias indicates how much the bootstrap estimate deviates from the original sample estimate. It is calculated as the difference between the average bootstrap estimate and the original estimate:

$$
\text{Bias}_{\text{boot}} = \frac{1}{B} \sum_{b=1}^B (\hat{\theta}^*_b - \hat{\theta})
$$

8. **Bootstrap Confidence Interval**
    - Confidence intervals for the statistic of interest can be constructed using the bootstrap distribution of $\hat{\theta}^*$. Several methods are explained below.


```{r}
# Bootstrapping
bootstrap_insect_data <- perform_bootstrap(
  data_cube_df = insect_data_new,
  fun = evenness_formula,
  samples = 1000,
  seed = 123)

# Summarise in dataframe
bootstrap_insect_data_full <- bootstrap_list_to_df(bootstrap_insect_data)
```

We get confidence intervals (`boot::boot.ci()`) and add it to the dataframe.
There are five types of intervals that can be calculated.
The default calculates all intervals.

Davison, A.C. & Hinkley, D.V. (1997). *Bootstrap Methods and Their Application*, Chapter 5. Cambridge University Press.

**1. Normal approximation** (normal)

Assuming the bootstrap distribution of the statistic is approximately normal.

$$
\left[ \hat{\theta} - \text{Bias}_{\text{boot}} - \text{SE}_{\text{boot}} \times z_{1-\alpha/2}, \hat{\theta} - \text{Bias}_{\text{boot}} + \text{SE}_{\text{boot}} \times z_{1-\alpha/2} \right]
$$

where $z_{1-\alpha/2}$ is the $1-\alpha/2$ quantile of the standard normal distribution for $\alpha$ confidence level.

**2. Basic bootstrap method** (basic)

Centering the interval around the original estimate $\hat{\theta}$.

$$
\left[ 2\hat{\theta} - \hat{\theta}^*_{(1-\alpha/2)}, 2\hat{\theta} - \hat{\theta}^*_{(\alpha/2)} \right]
$$

where $\hat{\theta}^*_{(\alpha/2)}$ and $\hat{\theta}^*_{(1-\alpha/2)}$ are the $\alpha/2$ and $1-\alpha/2$ percentiles of the bootstrap distribution, respectively.

**3. Studentized bootstrap method** (studentized)

Similar as **1**, but $z_{1-\alpha/2}$ estimated based on the bootstrap distribution.

**4. Bootstrap percentile method** (percent)

Using the percentiles of the bootstrap distribution.

$$
\left[ \hat{\theta}^*_{(\alpha/2)}, \hat{\theta}^*_{(1-\alpha/2)} \right]
$$

where $\hat{\theta}^*_{(\alpha/2)}$ and $\hat{\theta}^*_{(1-\alpha/2)}$ are the $\alpha/2$ and $1-\alpha/2$ percentiles of the bootstrap distribution, respectively.


**5. Adjusted bootstrap percentile method (Bias-Corrected and Accelerated)** (bca)

Adjusting for bias and skewness in the bootstrap distribution.

$$
\left[ \hat{\theta}^*_{(\alpha_1)}, \hat{\theta}^*_{(\alpha_2)} \right]
$$

where $\alpha_1$ and $\alpha_2$ are adjusted percentiles taking into account the bias and acceleration (adjust for skewness).

```{r}
# Calculate confidence intervals
ci_df <- get_bootstrap_ci(bootstrap_insect_data)

# Join dataframes
bootstrap_insect_data_final <- bootstrap_insect_data_full %>%
  full_join(ci_df, by = join_by(year), relationship = "many-to-many")
```

Calculation of studentized intervals failed.
Its use is further explained in `?boot::boot.ci()`.

## Visualise confidence intervals

We visualise the distribution of the bootstrap replicates.
The red dots are the the original sample estimates $\hat{\theta}$.

```{r}
bootstrap_insect_data_final %>%
  ggplot(aes(x = year)) +
    geom_boxplot(aes(y = est_boot)) +
    geom_point(aes(y = est_original), colour = "firebrick", size = 3) +
    labs(y = "evenness") +
    scale_y_continuous(limits = c(0, 1))
```

```{r}
bootstrap_insect_data_final %>%
  ggplot(aes(x = year)) +
    geom_violin(aes(y = est_boot)) +
    geom_point(aes(y = est_original), colour = "firebrick", size = 3) +
    labs(y = "evenness") +
    scale_y_continuous(limits = c(0, 1))
```

We visualise the confidence intervals.

```{r}
bootstrap_insect_data_final %>%
  distinct(year, est_original, se_boot, conf_name, ll, ul, conf_level) %>%
  ggplot(aes(x = year, y = est_original, group = conf_name)) +
    geom_errorbar(aes(ymin = ll, ymax = ul, colour = conf_name),
                  position = position_dodge(0.8), linewidth = 1) +
    geom_point(colour = "firebrick", size = 3, position = position_dodge(0.8)) +
    labs(y = "evenness") +
    scale_y_continuous(limits = c(NA, 1))
```

We compare the confidence intervals with the distribution of the bootstrap replicates.

```{r}
bootstrap_insect_data_final %>%
  ggplot(aes(x = year)) +
    geom_boxplot(aes(y = est_boot)) +
    geom_errorbar(aes(ymin = ll, ymax = ul, colour = conf_name),
                  position = position_dodge(0.8), linewidth = 1) +
    geom_point(aes(y = est_original), colour = "firebrick", size = 3) +
    scale_y_continuous(limits = c(NA, 1))
```

```{r}
bootstrap_insect_data_final %>%
  ggplot(aes(x = year)) +
    geom_violin(aes(y = est_boot)) +
    geom_errorbar(aes(ymin = ll, ymax = ul, colour = conf_name),
                  position = position_dodge(0.8), linewidth = 1) +
    geom_point(aes(y = est_original), colour = "firebrick", size = 3) +
    labs(y = "evenness") +
    scale_y_continuous(limits = c(NA, 1))
```
